Probing the inner kpc of massive galaxies with strong gravitational lensing

### 2015-01-07 response.txt

We are grateful to the referee for careful reading of the 
manuscript and helpful suggestions. We have addressed each of their 
points, highlighting the changes to the manuscript in blue.

@@

" a background source with a Gaussian light profile (FWHM 1170 pc)..."

==================================================================================

How was the black hole mass chosen? I assume it follows the M-sigma 
relation, but that is not discussed explicitly (perhaps because the 
effective velocity dispersion of the model galaxy is not 
mentioned).

>> RESPONSE: For a lens galaxy with a sigma of 200 km/s the black hole
>> mass on the M-sigma relation is about 2e8 Msun (from Kormendy & Ho 
>> 2013 and McConnell et al. 2011). The giant elliptical strong lenses 
>> have typical velocity dispersions of 200 km/s (Sonnenfeld et al. 
>> 2013), so we think that black holes in the 10^8 Msun mass range are 
>> plausible for these galaxies. We have added this information to the 
>> simulation section:

"Lensing galaxies are typically massive ellipticals with velocity 
dispersions of order 200 km/s (e.g., Sonnenfeld et al. 2013). From 
the M-sigma relation, we predict that these galaxies should contain 
black holes with masses in excess of 2 x 10^8 Msun (e.g., Kormendy & 
Ho 2013). In our simulations we have chosen black holes with masses 
of 1, 2, and 4 x 10^8 Msun."

@@

"The stellar population is modeled as an elliptical cored 
power-law. The deflection angles for elliptical cored power-law 
models were calculated using FastEll (Barkana 1998)."

==================================================================================

"For double and quad image configurations, the theory of strong 
lensing predicts a third and a fifth image near the centers of 
lensing galaxies." Strictly speaking, that statement applies only 
under certain conditions. Even if those conditions are thought to 
apply to realistic galaxy models, it is important to understand 
that the theoretical statement is nuanced. For example, there is no 
central image if the surface mass density has a central cusp that 
is steeper than 1/r, or if the SMBH exceeds a threshold in mass. If 
the black hole mass is below that threshold, there can actually be 
more than one central image.

>> RESPONSE: We have tried to qualify this with the sentences that
follow this description in the introduction:

### 2015-01-07 ms.tex
Tiny corrections, flagged changes in blue to referee.

\begin{abstract}  
\noindent
We examine the prospects of detecting demagnified images of gravitational lenses in observations of strongly lensed mm-wave molecular emission lines with ALMA. 
We model the lensing galaxies as a superposition of a dark matter component, a stellar component, and a central supermassive black hole and assess the detectability of the central images for a range of relevant parameters (e.g., stellar core and black hole mass).
We find that over a large range of plausible parameters, future deep observations of lensed molecular lines with ALMA should enable  detection of the central images at $\gtrsim 3\sigma$ significance. We use a Fisher analysis to examine the  constraints that could be placed on these parameters in various scenarios and find that for large stellar cores, both the core size and the mass of the central SMBHs can be accurately measured. We also study the prospects for detecting binary SMBHs with such observations and find that only under rare conditions and with very long integrations ($\sim$40-hr) the masses of both SMBHs may be measured using the distortions of central images.

@@

\section{introduction}
Probing the matter distribution in the innermost kpc of galaxies can answer key questions about super-massive black holes (SMBH), galaxy formation, and dark matter. It is now established that almost every massive galaxy harbors a SMBH at its center with a mass that strongly correlates with the mass of the host galaxy \citep{Kormendy:95,Ferrarese:00,Gebhardt:00,Tremaine:02}.
In addition, the distribution of stellar populations in central regions of galaxies contains information about their past merger histories and SMBH-stellar population interactions \citep[e.g.,][]{Barnes:92,Ebisuzaki:91,Barnabe:14}.  Various dark matter models also predict different structures for the central regions of dark matter halos \citep[e.g.,][]{Rocha:13}.  Mapping the matter density in the central regions of galaxies can thus shed light on various astrophysical phenomena.

@@

Strong gravitational lensing is a powerful tool for probing the matter distribution in distant galaxies. Among other things, strong lenses have been used to constrain galaxy masses, density profiles, and abundance of dark matter subhalos \citep[e.g.,][]{dalal:02, Gavazzi:07, Bolton:08}. For double and quad image configurations, the theory of strong lensing predicts a third and a fifth image near the centers of lensing galaxies. 
\blue{However, realistic galaxy density profiles produce central images that are almost always \emph{demagnified}\footnote{approximately as $(\Sigma/\Sigma_{crit})^{-2}$, where $\Sigma$ is the surface density and $\Sigma_{crit}$ is the critical surface density}, making their detection difficult \citep[e.g.,][]{Jackson:13}. If the lens galaxies have cuspy central profiles with steep slopes or very massive SMBHs, this central image may not form at all.}  
%\red{Unlike the other lensed images, which are magnified, this image can be significantly \emph{demagnified}, making its detection difficult \citep[e.g.,][]{Jackson:13}.}
 It is well-understood that the magnification of the central images is very sensitive to the matter distribution in the innermost regions of lens galaxies: very steep density profiles significantly demagnify the central images, whereas cored or shallow profiles render them brighter (Figure \ref{fig1}).

>>>>

\begin{abstract}

\noindent We examine the prospects of detecting demagnified images 
of gravitational lenses in observations of strongly lensed mm-wave 
molecular emission lines with ALMA. We model the lensing galaxies 
as a superposition of a dark matter component, a stellar component, 
and a central supermassive black hole and assess the detectability 
of the central images for a range of relevant parameters (e.g., 
stellar core and black hole mass). We find that over a large range 
of plausible parameters, future deep observations of lensed 
molecular lines with ALMA should enable detection of the central 
images at $\gtrsim 3\sigma$ significance. We use a Fisher analysis 
to examine the constraints that could be placed on these parameters 
in various scenarios and find that for large stellar cores, both 
the core size and the mass of the central SMBHs can be accurately 
measured. We also study the prospects for detecting binary SMBHs 
with such observations and find that only under rare conditions and 
with very long integrations ($\sim$40-hr) the masses of both SMBHs 
may be measured using the distortions of central images.

@@

\section{introduction}

Probing the matter distribution in the innermost kpc of galaxies can answer key questions about super-massive black holes (SMBH), galaxy formation, and dark matter. It is now established that almost every massive galaxy harbors a SMBH at its center with a mass that strongly correlates with the mass of the host galaxy \citep{Kormendy:95,Ferrarese:00,Gebhardt:00,Tremaine:02}.
In addition, the distribution of stellar populations in central regions of galaxies contains information about their past merger histories and SMBH-stellar population interactions \citep[e.g.,][]{Barnes:92,Ebisuzaki:91,Barnabe:14}.  Various dark matter models also predict different structures for the central regions of dark matter halos \citep[e.g.,][]{Rocha:13}.  Mapping the matter density in the central regions of galaxies can thus shed light on various astrophysical phenomena.

@@

Strong gravitational lensing is a powerful tool for probing the 
matter distribution in distant galaxies. Among other things, strong 
lenses have been used to constrain galaxy masses, density profiles, 
and abundance of dark matter subhalos \citep[e.g.,][]{dalal:02, 
Gavazzi:07, Bolton:08}. For double and quad image configurations, 
the theory of strong lensing predicts a third and a fifth image 
near the centers of lensing galaxies.
\blue{However, realistic galaxy density profiles produce central 
images that are almost always 
\emph{demagnified},\footnote{approximately as 
$(\Sigma/\Sigma_{crit})^{-2}$, where $\Sigma$ is the surface 
density and $\Sigma_{crit}$ is the critical surface density.} 
making their detection difficult \citep[e.g.,][]{Jackson:13}. If 
the lens galaxies have cuspy central profiles with steep slopes or 
very massive SMBHs, this central image may not form at all.}
%\red{Unlike the other lensed images, which are magnified, this image can be significantly \emph{demagnified}, making its detection difficult \citep[e.g.,][]{Jackson:13}.}
It is well-understood that the magnification of the central images 
is very sensitive to the matter distribution in the innermost 
regions of lens galaxies: very steep density profiles significantly 
demagnify the central images, whereas cored or shallow profiles 
render them brighter (Figure \ref{fig1}).
